### 2020-12-19 Thesis-Abstract.tex
Editorial tweaks

Public debate has resumed on the topic of exceptional access (EA), which refers to alternative means of decryption
intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of
the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
steps, it is necessary that the technical community remain engaged. Although any EA framework would increase risk, the
magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
EA may reduce risk overall.

The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a cornerstone of
security. The emergence of strong encryption incited the first crypto war, and its proliferation is causing the second.
In response to both polarized and conciliatory voices, this paper analyzes strategies for confronting wicked problems
and proposes an iterative approach to the case of encryption and EA. Along the way, it illustrates the components of the
debate in argument maps and demonstrate the security risks with data flow diagrams and threat analysis, focusing on one
EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk.''

>>>>

\parboxindent Public debate has resumed on the topic of exceptional access (EA), which refers to alternative means of
decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
steps, it is necessary that the technical community remain engaged. Although any EA framework would increase risk, the
magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
EA may reduce risk overall.

\parboxindent The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a
cornerstone of security. The emergence of strong encryption incited the first crypto war, and its proliferation is
causing the second. In response to both polarized and conciliatory voices, this paper analyzes strategies for
confronting wicked problems and proposes an iterative approach to the case of encryption and EA. Along the way, it
illustrates the components of the debate in argument maps and demonstrate the security risks with data flow diagrams and
threat analysis, focusing on one EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass
Surveillance Risk.''

### 2020-12-19 Thesis-Chapter2-Background.tex
Editorial tweaks

In 1946 The U.S. formalized an intelligence-sharing operation with the U.K. in the BRUSA (now UKUSA) Agreement. Over the
following 10 years, Australia, Canada, and New Zealand were made full partners to this agreement, which has become known
as ``Five Eyes'' \cite{nsa_2020}. Due to the close relationship between the intelligence and law enforcement agencies of
these countries, the member's encryption policies are relevant to each another.\\

@@

The U.K. government has the explicit ability to force decryption, perform \ac{lawful-hacking}, and mandate \ac{EA}
though ``technical capability notices'' introduced in 2016's Investigatory Powers Act \cite{legislature_2016}. Australia
has weaker powers through a variety of laws, and in 2018 passed the Assistance and Access Act which created very similar
``technical capability notices'' \cite{ag_2018}. Canadian law requires cooperation between telecommunications providers
and law enforcement but does not directly address encryption. Canada has traditionally supported strong encryption, but
the pressure of \ac{the-cw2} is straining that support \cite{parsons_2019}.

>>>>

In 1946 The U.S. formalized an intelligence-sharing operation with the U.K. in the BRUSA (now UKUSA) Agreement. Over the
following 10 years, Australia, Canada, and New Zealand were made full partners to this agreement, which has become known
as ``Five Eyes'' \cite{nsa_2020}. Due to the close relationship between the intelligence and law enforcement agencies of
these countries, the member's encryption policies are relevant to each another.

@@

\parboxindent The U.K. government has the explicit ability to force decryption, perform \ac{lawful-hacking}, and mandate
\ac{EA} though ``technical capability notices'' introduced in 2016's Investigatory Powers Act \cite{legislature_2016}.
Australia has weaker powers through a variety of laws, and in 2018 passed the Assistance and Access Act which created
very similar ``technical capability notices'' \cite{ag_2018}. Canadian law requires cooperation between
telecommunications providers and law enforcement but does not directly address encryption. Canada has traditionally
supported strong encryption, but the pressure of \ac{the-cw2} is straining that support \cite{parsons_2019}.

### 2020-12-19 Thesis-Chapter3-PolicyMaking.tex
Update forms, small editorial tweaks
Editorial tweaks

Cybersecurity law scholar Alan Rozenshtein argues at length for treating \ac{encryption} and \ac{EA} as a
\ac{wicked-prob} \cite{rozenshtein_wicked_2018}. He divides the root issues into three categories. First, there is
disagreement on what the goals should be (and even basic premises). Sides do not agree on how much to value competing
notions of security. Regarding the basic facts, sides do not agree on whether encryption is hiding so much evidence that
law enforcement is ``going dark,'' as one side puts it, or whether technological change overall has created a ``golden
age for surveillance,'' as their opponents argue. Second, information is ``uncertain and diffuse.'' Comprehensive data
regarding encryption's effect on investigations is unavailable. Although the consensus is that we are not presently
capable of acceptably secure \ac{EA}, that consensus could change, especially considering that \ac{EA} as a field is
under-researched. Third, that the problem cannot be definitively solved. Evolving values and technology mean that this
policy area is always up for renegotiation.

@@

\Ac{incrementalism} has several advantages over the \ac{classical-method} for handling \acp{wicked-prob}. It is rooted
in realism about the limits of rational analysis. It accepts that the problem will not be conclusively solved. Instead,
it emphasizes iteration: ``Policy is not made once and for all; it is made and re-made endlessly. Policy-making is a
process of successive approximation to some desired objectives in which what is desired itself continues to change under
reconsideration'' \cite{lindblom_muddling_1959}. Most importantly, it eschews the linear
problem-definition--alternatives-analysis steps out of respect that the problem definition is not self-contained.
Lindblom argues that policy ends and means are interlinked, eventually concluding:

@@

The \acl{OODAloop} was developed by Air Force Colonel John Boyd as a description of a successful strategy for countering
opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but its ideas
have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring incident
response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your opponent's
loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.

@@

The \ac{OODAloop} serves well as a model for contending with \acp{wicked-prob} because it has the correct fundamental
structure. It corresponds closely to the \ac{classical-method} steps of problem definition (orientation), ideation and
analysis (orientation), decision, and action; however, it also emphasizes \ac{incrementalism}'s iteration and feedback.
Additionally, while policymaking does not share the pace of real-time combat, this metaphor holds true in other ways. In
both cases, one faces an unpredictable opponent in a dynamic environment in which every action counts. We cannot
overpower \acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.

@@

\begin{figure}[h]
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{dfds/build/OODA-policy-process.png}
  \caption{Using the \ac{OODAloop} to Tackle Wicked Problems}
  \label{fig-policy-ooda-process}
\end{figure}

>>>>

Cybersecurity law scholar Alan Rozenshtein argues at length for treating \ac{encryption} and \ac{EA} as a
\ac{wicked-prob} \cite{rozenshtein_wicked_2018}. He divides the root issues into three categories. First, there is
disagreement on what the goals should be (and even basic premises). Sides do not agree on how much to value competing
notions of security. Regarding the basic facts, sides do not agree on whether encryption is hiding so much evidence that
law enforcement is ``going dark,'' as one side puts it, or whether technological change overall has created a ``golden
age for surveillance,'' as their opponents argue. Second, information is ``uncertain and diffuse.'' Comprehensive data
regarding encryption's effect on investigations is unavailable. Although the consensus is that we are not presently
capable of acceptably secure \ac{EA}, that consensus could change, especially considering that \ac{EA} as a field is
under-researched. Third, the problem cannot be definitively solved. Evolving values and technology mean that this policy
area is always up for renegotiation.

@@

\Ac{incrementalism} has several advantages over the \ac{classical-method} for handling \acp{wicked-prob}. It is rooted
in realism about the limits of rational analysis. It accepts that the problem will not be conclusively solved. Instead,
it emphasizes iteration: ``Policy is not made once and for all; it is made and re-made endlessly. Policy-making is a
process of successive approximation to some desired objectives in which what is desired itself continues to change under
reconsideration'' \cite{lindblom_muddling_1959}. Most importantly, it eschews the linear flow from problem definition to
alternatives analysis out of respect that the problem definition is not self-contained. Lindblom argues that policy ends
and means are interlinked, eventually concluding:

@@

The \ac{OODAloop} was developed by Air Force Colonel John Boyd as a description of a successful strategy for countering
opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but its ideas
have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring incident
response \cite{schneier_future_2014}. The model emphasizes fast cycle times and ``getting inside'' your opponent's loop
as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.

@@

The \ac{OODAloop} serves well as a model for contending with \acp{wicked-prob} because it has the correct fundamental
structure. It corresponds closely to the \ac{classical-method} steps of problem definition (orientation), ideation and
analysis (orientation), decision, and action; however, it also emphasizes \ac{incrementalism}'s iteration and feedback.
Additionally, while policymaking does not share the pace of real-time combat, this metaphor holds true in other ways. In
both cases, one faces an unpredictable opponent in a dynamic environment in which every action counts. We cannot
overpower \acp{wicked-prob} through force of reason, but perhaps we can through feedback and agility.

@@

\begin{figure}[h]
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{dfds/build/OODA-policy-process.png}
  \caption{Using the \acs{OODAloop} to Tackle Wicked Problems}
  \label{fig-policy-ooda-process}
\end{figure}

### 2020-12-19 Thesis-Chapter4-Tools.tex
Update forms, small editorial tweaks
Editorial tweaks

In order to advance the \ac{EA} discussion, it is necessary to organize the many different arguments in a cohesive
manner. This thesis uses \acp{argmap} to do so. \Acp{argmap} are graphical representations of the logical structures and
relationships between statements, premises, and conclusions. They originated from the same research as the idea of
\acp{wicked-prob}. Though formal research on the technique is sparse, \acp{argmap} have proven helpful in simplifying
complex arguments regarding \acp{wicked-prob} \cite{renton_2007} and in facilitating debate. This is because they slow
down discussion, depersonalize conflict, and lend structure and rhythm to meetings \cite{dutoit_hypermedia_2006}. They
are therefore useful as a tool for conducting the \ac{OODAloop} model's ``collaborative debate'' step.

@@

\begin{figure}[h]
    \centering\CaptionFontSize
    \includegraphics[width=0.8\linewidth]{dfds/build/diffie-hellman-traditional.png}
    \caption{A \acf{DFD} for a simplified Diffie-Hellman key exchange}
    \label{fig-dfd-dh-traditional}
\end{figure}

>>>>

In order to advance the \ac{EA} discussion, it is necessary to organize the many different arguments in a cohesive
manner. This thesis uses \acp{argmap} to do so. \Acp{argmap} are graphical representations of the logical structures and
relationships between statements, premises, and conclusions. They originated from the same research as the idea of
\acp{wicked-prob}. Though formal research on the technique is sparse, \acp{argmap} have proven helpful in simplifying
complex arguments regarding \acp{wicked-prob} and in facilitating debate \cite{renton_2007}. This is because they slow
down discussion, depersonalize conflict, and lend structure and rhythm to meetings \cite{dutoit_hypermedia_2006}. They
are therefore useful as a tool for conducting the \ac{OODAloop} model's ``collaborative debate'' step.

@@

\begin{figure}[h]
    \centering\CaptionFontSize
    \includegraphics[width=0.8\linewidth]{dfds/build/diffie-hellman-traditional.png}
    \caption{A \acs{DFD} for a simplified Diffie-Hellman key exchange}
    \label{fig-dfd-dh-traditional}
\end{figure}

### 2020-12-19 Thesis-Chapter5-Arguments.tex
Editorial improvements
Editorial tweaks

This chapter will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
as tools in the collaborative pursuit of the problem definition and solution. It is not a replacement for debate arising
from live discussion. Second, \acp{argmap} may deceptively portray the strength of an argument. Argument nodes do not
indicate their strength or validity. If ten unsound arguments were presented against one indisputable argument, the side
with more arguments would, regardless of their integrity, appear stronger. With those qualifications in mind, let us
begin by examining the factors at the center of the conflict.

@@

The ``Going Dark'' argument has a few shortcomings. It oversimplifies the relationship between privacy and safety by
depicting it as a zero-sum conflict. In reality, privacy does not always diminish safety, but sometimes enhances safety.
For example, data privacy makes crimes such as stalking and identify more difficult. The ``Going Dark'' argument also
categorizes privacy as a personal good and safety as a public good. However, privacy can also be a public good when it
precludes \ac{masssurv}. This argument is explored more deeply in the ``Golden Age for Surveillance'' map.

@@

However, lack of evidence for an argument does not prove that it is incorrect. Because law enforcement has the data, the
burden of proof is on them---they must show that encryption is causing investigatory failure. Unfortunately, in most
cases, it is impossible to know what would have happened without the interference of strong encryption. Inability to
gather data from controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable,
another way to gather data is by observing naturally-occurring conditions in which everything but the variable of
interest remain are constant. These conditions exist for encryption in messaging application ecosystem. Messaging
applications WhatsApp, Facebook Messenger, and Telegram all have large user bases and similar features. Among these
services, only Facebook Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market
share as measured by active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports
\cite{keller_internet_2019}. This indicates that encryption does indeed affect investigations.

@@

Encryption is not the only challenge that law enforcement faces with digital evidence. A 2018 study found that
identifying service providers and getting data from them using existing mechanisms were even greater challenges
\cite{carter_2018}. The report, titled ``Low-Hanging Fruit,'' also lists resource limitations and poor training as top
issues. Indeed, many current initiatives are underfunded. The \ii{New York Times}' 2019 investigation into \ac{CSAM}
uncovered a chronic lack of commitment from government initiatives \cite{keller_internet_2019}. Major legislation passed
in 2008, but has been funded at only 50\% of authorized levels; a commissioned Justice Department task force has
produced only two of five biennial reports; and a senior executive position within the Justice Department to target
\ac{CSAM} was never created. Federal entities aiming to improve training and coordination between law enforcement and
industry are similarly underfunded \cite{carter_2018}.

@@

It is tempting to declare government requests for \ac{EA} disingenuous due to their hypocritical behavior towards some
of these issues. For example, if combatting \ac{CSAM} is a government priority, why has Congress underfunded the laws
it's passed, and why has the Justice Department understaffed their its task force \cite{keller_internet_2019}? These
questions demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The absence of
government funding does not mean technical anti-\ac{CSAM} measures are unnecessary.

@@

In \myfig{fig-arg-measures}, response measures are considered according to five metrics. These metrics are based on
several sources, including two separate diverse committees gathered to analyze the encryption debate
\cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
of security, safety, privacy, and trust. The five metrics used to analyze response measures are (1) minimization of
increased security risk, (2) utility to law enforcement, (3) protection of civil liberties---particularly important as
after the analysis of the ``Golden Age for Surveillance'' argument---(4) transparency, and (5) economic impact. Economic
impact is not as central an issue as the others, but it does help determine practical viability of a proposal.

@@

We could also decide to increase investment in current technologies and investigative techniques. As previously
discussed, encryption is not the most significant barrier in law enforcement's use of digital evidence
\cite{carter_2018}. Even when encryption is involved, traditional sleuthing can often lead to access. Security expert
Bruce Schneier and cyberlaw expert Orin Kerr have compiled a list of ``encryption workarounds'' that do not rely on
\ac{EA}. These include finding the key, guessing the key, accessing plaintext while the device is in use, and finding
another copy of the plaintext \cite{kerr_encryption_2017}. Though probabilistic and time-consuming in comparison to
\ac{EA}, these methods do not require new legal or technical capability. These methods are effective
\cite{greenberg_2018}; investment in current digital evidence capabilities and anti-\ac{CSAM} task forces, they are also
safe.

@@

\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA} \cite{bellovin_lawful_2013}
\cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018} \cite{kerr_encryption_2017} \cite{soesanto_2018}.
\Ac{lawful-hacking}'s support may be due in part to the fact that it is a technical approach that isn't \ac{EA} itself.
However, its effectiveness against criminals for whom \ac{EA} would not work is its obvious benefit. This is because,
while organized crime and terror groups would still use encryption tools to evade an \ac{EA} scheme, their operational
security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
This approach has already been used \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
\cite{feuer_chapo_2019}. Because it is already used, using \ac{lawful-hacking} as the primary strategy would require
officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.

@@

Trusted-party key escrow relies on an entity or a small group of entities to store a key or key recovery information.
There were a wide variety of trusted-party key escrow implementations created during \ac{the-cw1}
\cite{denning_taxonomy_1996}. This type of \ac{EA} has a crushing weakness: it centralizes extremely sensitive
information, which heightens the risk of a disastrous data breach. Some argue software companies are capable of
protecting such sensitive data, citing their current practice of storing code-signing keys in \acp{HSM}
\cite{ozzie_2018}. However, such keys are rarely used in comparison to \ac{EA} keys, which would be updated and accessed
frequently. Although code-signing keys are rarely used and are protected in \acp{HSM}, they still leak
\cite{green_2018}. Once the escrowed keys leak, every device or message encrypted with them would be at risk. This is an
example of the \ac{argmap}'s sometimes imbalanced portrayal of an argument's strength---in this case, the map
underrepresents the effectiveness of the security-based argument against centralized access mechanisms.

@@

One type of \ac{DIM} \ac{EA} involves uses strong cryptography but includes extra cryptographic ``puzzle pieces'' in
message metadata. A surveiller could reconstruct the encryption key from the puzzle pieces, but only after expending
considerable computational power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes mass
surveillance, but it also enables arbitrary surveillance by anyone with enough computing power. Using this type of
\ac{DIM} \ac{EA} in common cases would be impractical due to its high cost. However, it would be less effective in
high-profile cases due to the ability of sophisticated adversaries to employ \ac{EA} workarounds. Additionally, puzzle
difficulty would be calibrated based on the changing standards of computational capability. Since opponents can record
data and decrypt it later when capabilities have changed, the data becomes steadily less secure over time.

Another \ac{DIM} \ac{EA} proposal involves adding authorities as a ``ghost'' participant in a conversation, analogous to
a silent listener on a conference call. This was notably suggested by \ac{GCHQ} officials Ian Levy and Crispin Robinson
\cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:

@@

Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
lessons of \mychap{chap-policy}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has
led to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This
is because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
privacy and civil liberties more significantly than criminal abuse.

>>>>

This chapter will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are not a
replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively portray the strength of an
argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments were presented against one
indisputable argument, the side with more arguments would, regardless of the arguments' integrity, appear stronger. With
those qualifications in mind, let us begin by examining the factors at the center of the conflict.

@@

The ``Going Dark'' argument has a few shortcomings. It oversimplifies the relationship between privacy and safety by
depicting it as a zero-sum conflict. In reality, privacy does not always diminish safety, but sometimes enhances safety.
For example, data privacy makes crimes such as stalking and identity theft more difficult. The ``Going Dark'' argument
also categorizes privacy as a personal good and safety as a public good. However, privacy can also be a public good when
it counters \ac{masssurv}. This argument is explored more deeply in the ``Golden Age for Surveillance'' map.

@@

However, lack of evidence for an argument does not prove that it is incorrect. Because law enforcement has the data, the
burden of proof is on them---they must show that encryption is causing investigatory failure. Unfortunately, in most
cases, it is impossible to know what would have happened without the interference of encryption. Inability to gather
data from controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable, another
way to gather data is by observing naturally-occurring conditions in which everything but the variable of interest are
constant. These conditions exist for encryption in the messaging application ecosystem. Messaging applications WhatsApp,
Facebook Messenger, and Telegram all have large user bases and similar features. Among these services, only Facebook
Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market share as measured by
active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports \cite{keller_internet_2019}. This
indicates that encryption does indeed affect investigations.

@@

Encryption is not the only challenge that law enforcement faces with digital evidence. A 2018 study found that
identifying and cooperating with service providers through existing legal frameworks were even greater challenges
\cite{carter_2018}. The report, titled ``Low-Hanging Fruit,'' also lists resource limitations and poor training as top
issues. Indeed, many current initiatives are underfunded. The \ii{New York Times}' 2019 investigation into \ac{CSAM}
uncovered a chronic lack of commitment from government initiatives \cite{keller_internet_2019}. Major legislation passed
in 2008, but has been funded at only 50\% of authorized levels; a commissioned Justice Department task force has
produced only two of five biennial reports; and a senior executive position within the Justice Department to target
\ac{CSAM} was never created. Federal entities aiming to improve training and coordination between law enforcement and
industry are similarly underfunded \cite{carter_2018}.

@@

It is tempting to declare government requests for \ac{EA} disingenuous due to their hypocritical behavior towards some
of these issues. For example, if combatting \ac{CSAM} is a government priority, why has Congress underfunded the laws
it's passed, and why has the Justice Department understaffed its task force \cite{keller_internet_2019}? These questions
demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The absence of government
funding does not mean technical anti-\ac{CSAM} measures are unnecessary.

@@

In \myfig{fig-arg-measures}, response measures are considered according to five metrics. These metrics are based on
several sources, including two separate diverse committees gathered to analyze the encryption debate
\cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
of security, safety, privacy, and trust. The five metrics used to analyze response measures are (1) minimization of
increased security risk, (2) utility to law enforcement, (3) protection of civil liberties---particularly important as
after the analysis of the ``Golden Age for Surveillance'' argument---(4) transparency, and (5) economic impact. Economic
impact is not as central an issue as the others, but it does affect the practical viability of a proposal.

@@

We could also decide to increase investment in current technologies and investigative techniques. As previously
discussed, encryption is not the most significant barrier in law enforcement's use of digital evidence
\cite{carter_2018}. Even when encryption is involved, traditional sleuthing can often lead to access. Security expert
Bruce Schneier and cyberlaw expert Orin Kerr have compiled a list of ``encryption workarounds'' that do not rely on
\ac{EA}. These include finding the key, guessing the key, accessing plaintext while the device is in use, and finding
another copy of the plaintext \cite{kerr_encryption_2017}. Though probabilistic and time-consuming in comparison to
\ac{EA}, these methods do not require new legal or technical capability. These methods are effective
\cite{greenberg_2018}; due to their conservative nature, they are also safe.

@@

\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA} \cite{bellovin_lawful_2013}
\cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018} \cite{kerr_encryption_2017} \cite{soesanto_2018}.
\Ac{lawful-hacking}'s support may be due in part to the fact that it is a technical approach that isn't \ac{EA} itself.
However, its effectiveness against criminals for whom \ac{EA} would not work is its obvious benefit. This is because,
while organized crime and terror groups would still use encryption tools to evade an \ac{EA} scheme, their operational
security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
This approach has already been used \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
\cite{feuer_chapo_2019}. Because it is already used, choosing \ac{lawful-hacking} as the primary strategy would require
officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.

@@

Trusted-party key escrow relies on an entity or a small group of entities to store a key or key recovery information.
There were a wide variety of trusted-party key escrow implementations created during \ac{the-cw1}
\cite{denning_taxonomy_1996}. This type of \ac{EA} has a crushing weakness: it centralizes extremely sensitive
information, which heightens the risk of a disastrous data breach. Some argue software companies are capable of
protecting such sensitive data, citing their current practice of storing code-signing keys in \acp{HSM}
\cite{ozzie_2018}. However, such keys are rarely used in comparison to \ac{EA} keys, which would be updated and accessed
frequently. Additionally, although code-signing keys are rarely used and are protected in \acp{HSM}, they still leak
\cite{green_2018}. Once the escrowed keys leak, every device or message encrypted with them would be at risk. This is an
example of the \ac{argmap}'s sometimes imbalanced portrayal of an argument's strength---in this case, the map
underrepresents the effectiveness of the security-based argument against centralized access mechanisms.

@@

One type of \ac{DIM} \ac{EA} uses strong cryptography but includes extra cryptographic ``puzzle pieces'' in message
metadata. A surveiller could reconstruct the encryption key from the puzzle pieces, but only after expending
considerable computational power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes mass
surveillance, but it also enables arbitrary surveillance by anyone with enough computing power. Using this type of
\ac{DIM} \ac{EA} in common cases would be impractical due to its high cost. However, it would be less effective in
high-profile cases due to the ability of sophisticated adversaries to employ \ac{EA} workarounds. Additionally, puzzle
difficulty would be calibrated based on the changing standards of computational capability. Since opponents can record
data and decrypt it later when capabilities have changed, the data becomes steadily less secure over time.

Another \ac{DIM} \ac{EA} proposal involves adding authorities as a ``ghost'' participant in a conversation, analogous to
a silent listener on a conference call. This was notably suggested by UK \ac{GCHQ} officials Ian Levy and Crispin
Robinson \cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:

@@

Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
lessons of \mychap{chap-policy}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has
led to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This
is because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
privacy and civil liberties more systematically than criminal abuse.
